### 2020-03-30 docs/draft_chapter.Rmd
start

---
title: "Topic 6.2: Open Data"
author: "Samuel Langton and Reka Solymosi"
date: "1 June 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

- Topic definition
- Chapter structure

## Background

- What is open data?
- What types of organisations release open data?
  - Public sector
  - Private sector
- What are the strengths/limitations of open data?

## Surveying environmental features using Open Street Map

- Explain OSM
- Note important technical details including key, value and elements
- Explain APIs
- Overpass queries and overpass-turbo
- But better: use OSM API to download some features in R
- Note the timeout error
- Map features
- Download crime data to compare with features (eg violence ~ bars)

## Future of open data
- Open topics in open data
- Further applications to crime and place research

## Conclusion

- Re-cap
- Wrap-up